---
title: "ESM 244 Assignment 2 - Task 2"
author: "Lucas Empson"
date: "2/8/2021"
output: 
  html_document:
    theme: darkly
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(GGally)
library(feasts)
library(broom)
library(jtools)
library(knitr)
library(kableExtra)
```

# Binary logistic regression to predict palmetto species

Use binary logistic regression to test feasibility of using variables 

- plant height (height) 
- canopy length (length)
- canopy width (width)
- number of green leaves (green_lvs) 

to classify whether a palmetto is species Serenoa repens or Sabel etonia. 

1: Serenoa repens
2: Sabal etonia

Use code folding and hide all messages & warnings in your knitted HTML.

```{r}
# Read in palmetto data, keeping only species, height, length, width, and green_lvs variables
palmetto <- read_csv(here("data", "palmetto.csv")) %>%
  clean_names() %>% 
  select(species, height, length, width, green_lvs) %>% 
  mutate(species = case_when(
    species == 1 ~ "Serenoa repens", # Change numerical observations to their corresponding species names
    species == 2 ~ "Sabal etonia")
    ) %>% 
  mutate(species = as_factor(species)) # Convert to factor

# Factor levels: Serenoa repens will be 0, Sabal etonia will be 1
# levels(palmetto$species)
```

# Main Content {.tabset}
## Tab 1: Simple visualizations of trends across variables 
```{r}
# Data viz of green_lvs (x-axis) and height (y-axis) by species (color)
ggplot(data = palmetto, aes(x = green_lvs, y = height)) +
  geom_point(aes(color = species)) +
  labs(x = "Number of green leaves", y = "Height") +
  theme_minimal() +
  scale_color_manual(values=c("#E1BE6A", "#40B0A6"))

# Data viz of length (x-axis) and height (y-axis) by species (color)
ggplot(data = palmetto, aes(x = length, y = height)) +
  geom_point(aes(color = species)) +
  labs(x = "Length", y = "Height") +
  theme_minimal() +
  scale_color_manual(values=c("#E1BE6A", "#40B0A6"))
  
# Data viz of width (x-axis) and length (y-axis) by species (color)
ggplot(data = palmetto, aes(x = width, y = length)) +
  geom_point(aes(color = species)) +
  labs(x = "Width", y = "Height") +
  theme_minimal() +
  scale_color_manual(values=c("#E1BE6A", "#40B0A6"))
```

## Tab 2: Binary logistic regression (BLR)
BLR using the four variables (height, length, width, & green_lvs) to predict which species
```{r, comment = NA}
# glm
palmetto_blr <- glm(species ~ height + length + width + green_lvs,
                    data = palmetto,
                    family = "binomial")

# BLR results:
palmetto_blr
summary(palmetto_blr)

# Get tidy version of the BLR results using `broom::tidy()`
palmetto_blr_tidy <- broom::tidy(palmetto_blr) # All coefficients are significant -- ****Why is `green_lvs` coefficient's p-value exactly 0?***
```

```{r}
# Convert log odds to probability: 
palmetto_blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>% # `.fitted` column shows probability of the palmetto species being "Sabel etonia" based on the four variables `height`, `length`, `width`, and `green_lvs`
  mutate(predicted_species = ifelse(.fitted < 0.5, "Serenoa repens",
                                    ifelse(.fitted >= 0.50, "Sabal etonia",
                                           "NA")))

```

## Tab 3: Model classification verification 
Finalized table showing, for each species, how many plants in the original dataset would be correctly classified and how many were incorrectly classified by the model

```{r}
table_stats <- palmetto_blr_fitted %>% 
  group_by(species, predicted_species) %>% 
  tally()

final_table <- data.frame("Species" = c("Serenoa repens","Sabal etonia"), 
           "Correct" = c(5548,5701), "Incorrect" = c(564,454),
           "Percent_Correctly_Classified"= c(90.77,92.62))

final_table %>% 
  kbl(format = "html",
      align = "lccc",
      col.names = c("Species", "Correct", "Incorrect", "% Correctly Classified"),
      caption = "Table 1. Model classification success summary. Number of times the binomial logistic regression correctly classified each species based on the explanatory variables.") %>%
  kable_styling(c("hover","condensed")) %>% 
  column_spec(1, italic = T)
  
```




## Tab 4: Visualizations of p("Sabel etonia")
Probability of being "Sabel etonia" by predictor variable using `jtools::effect_plot()`
```{r}
# height
effect_plot(palmetto_blr,
            pred = height,
            interval = TRUE,
            y.label = "Probability of being `Sabal etonia`")

# length
effect_plot(palmetto_blr,
            pred = length,
            interval = TRUE,
            y.label = "Probability of being `Sabal etonia`")
# width
effect_plot(palmetto_blr,
            pred = width,
            interval = TRUE,
            y.label = "Probability of being `Sabal etonia`")

# green_lvs
effect_plot(palmetto_blr,
            pred = green_lvs,
            interval = TRUE,
            y.label = "Probability of being `Sabal etonia`")
```













